### 2020-01-03 docs/thesis-project/article.tex
Update article.tex

\begin{itemize}
    \item Identify a steady-state, $Q$, providing optimal robustness and load-balancing
    \item Identify a transition matrix, $M$, that allows peers to achieve $Q$ collectively
    \item Optimize $M$ with respect to peer connectivity degree
    \item Optimize updates to $M$ as overlay topology changes over time
    \item Implement a replication model that exploits the properties of the $M$
    \item Evaluate solution performance in a Simulation or on in real life scenario
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
% END INTRODUCTION %
%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{images/gfs_paper.png}
\caption{GFS Architecture}
\label{fig:gfs_arch}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.6\textwidth]{images/cephfs.png}
\caption{CFS Architecture}
\label{fig:cfs_arch}
\end{figure}

@@

\section{P2P Overlays}\label{sec:p2p-overlays}

\begin{figure}[t]
    \centering
    \subfigure{\label{fig:iplayer}\includegraphics[width=0.35\textwidth, clip=true]{images/ip-layer.png}}
    \hspace{0.2\textwidth}
    \subfigure{\label{fig:overlay}\includegraphics[width=0.35\textwidth, clip=true]{images/overlay-layer.png}}
    \caption{P2P Network Diagrams: IP layer \textit{vs.} One possible overlay abstraction}
    \label{fig:network_diagrams}
\end{figure}

@@

\section{Membership Management Frameworks}\label{sec:membership-management}
Gossip-based protocols are well known for their usability to solve problems like database replication, failure detection, and resource-monitoring. We present two frameworks used for membership management that ensure peers keep up to date partial views of large-scale, dynamic networks, built on unstructured overlays using such protocols. Just like we mentioned in section \ref{sec:p2p-overlays}, keeping a view of all peers in the system is often not a solution. Another fundamental issue in P2P is how to avoid partitioning. The following alternatives are both scalable, robust, and decentralized solutions to both these problems. They gift peers with autonomous capabilities regarding self-organization and overlay healing.

@@

\section{Cloud-Assisted P2P Networks}\label{sec:ca-networks}

@@

\section{Data Redundancy}\label{sec:dataredundancy}
In DFS, not losing data is of the utmost importance. Particularly, peer availability in P2P networks is ever-changing and not guaranteed, and thus, availability, and even durability of files are at risk. A \textbf{\textit{durable file}} is one that, once put into the system, is not lost due to failures, whereas availability means that a user can access it promptly. Durability is accomplished by employing file or block-level \textbf{\textit{replication}} or alternatively, using \textbf{\textit{erasure coding}} techniques (\textit{coding} for short). Hybrid approaches are also possible. Replication has the advantage of having a straightforward and less resource-intensive recovery model, on the other hand, coding has a better \textit{expansion factor}, i.e., can provide the same or even increased durability at a lower storage overhead. Because our proposal focuses on durability of files in P2P networks built on top of personal computers, where resources may be limited, we survey only coding, mentioning raised concerns and inherent disadvantages of this technique.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN SOLUTION PROPOSAL %
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Solution Proposal}\label{sec:proposal}
In this section, we introduce a prototype architecture that targets the use case that inspired to explore the theme of this thesis. We explain it by decomposition.

\subsection{Use Case}
Fundamentally, the idea is that a company, Hives, with dedicated web servers called Hiveminds, provides a platform for registered users to sell storage belonging to their personal computers directly to other registered users. Clients pay a monthly fee directly to Hives based on how much storage space they need. Hiveminds then put together a set of sellers, which constitute a Hive, that serves adequate storage space to the clients. Sellers are selected according to the reputation they earn over time based on their reliability, with a bias towards more reputable sellers. The collected fee is distributed proportionately among Hive members according to their overall space contribution and up time, calculated weekly, for example. The company's revenue comes from keeping a small percentage of the monthly fees. If the client requires continuous availability, rather than durability, he may pay a premium, in which case, inspired on the related work mentioned in section \ref{sec:ca-networks}, cloud storage may be used to help the Hive in meeting the required service guarantees. Hive revenue shares are reduced if their reliability degrades too much.

@@

\subsection{Hiveminds - Master Servers}
The Hivemind servers are the backbone of the system; they have to handle registration, login, subscription and payment procedures. When a client wishes to subscribe to a monthly plan, he contacts one Hivemind server at random by accessing his desktop application, and the Hivemind then sets a \textit{usedSpace} variable indicating how much data the Client uploaded into the system to zero. This variable is updated according to the Client's future requests. In Hives File System (Hives), users' files are uniquely identified and have no hierarchy. Hiveminds keep a mapping between the \textit{fileId} and the Hive\footnote{Set of peers S responsible for holding one or more files from one or more clients} that is responsible for storing them. Hive membership is dynamic, and a mapping between the \textit{HiveId} and the Hive members (Workers) is also kept. Note that it is not relevant to know which fragments each Worker in the Hive has; this is important because it diminishes the number of messages sent to Hiveminds and also reduces space overhead required to keep track of file locations. When a Hivemind receives a file upload request from a Client, it finds or creates a suitable Hive to host the data and calculates, if needed, the transition matrix used for PSG within it, before returning all relevant data to the Client, including Workers' addresses. When the Hivemind receives a delete request, it drops all mapping data regarding the respective \textit{fileId}.

\subsection{Clients}
After registration, logging in, and subscription procedures, Clients, can perform four main procedures, which we name and implement using HTTP semantics. In all cases, the Client first retrieves the Hive Workers' addresses. In a GET request, the Client downloads enough fragments from the Hive Workers and reconstructs the file himself. In a DELETE request, the Client first asks the Hivemind to stop tracking efforts, and once he receives an acknowledge, he broadcasts in a best-effort manner a DELETE message to any online worker within that file's Hive. In a POST operation, the Client first obtains permission to upload from the Hivemind; after that, he divides his file into parts, applies an erasure coding algorithm over each them, and uploads the encoded fragments to the workers at random. PUT updates existing files and operate similarly to POST, but whenever possible, try to send the new version to the Hive that contains the previous \textit{versionId}.

\subsection{Workers - Sellers}
Apart from adequately dealing with clients' requests, Workers organize themselves in a P2P fashion within each Hive. One Worker can be in zero or more Hives, and within each Hive, they continuously redistribute encoded file part fragments according to the transition matrix that is calculated by the Hivemind and delivered to them by a Client. In the long run, we expect that a Hive achieves the desired fragment distribution that fairly distributes load while ensuring optimal durability. Constantly exchanging fragments is expected to cause a bandwidth overhead but has the advantage that Clients can upload their fragments to any online peer or peers in the Hive instead of one single entity that may go offline during the upload. Workers will then redistribute the fragment using simple routing rules. This approach does not sacrifice fast lookup operations due to the structured nature of the system using a Hivemind that behaves somewhat like a BitTorrent tracker. The never-ending fragments exchange between peers also facilitates the timely detection of non-transient faults in the Hive, which helps to ensure file durability. Workers piggyback aggregate information to the fragments, including but not limited to, how many parts of a file they have, allowing us to implement a hybrid between a reactive and a proactive system concerning replica control, in hopes of minimizing bandwidth burst on top of already high bandwidth requirements. Fragment exchange may eventually be replaced with ping-like messages to minimize bandwidth cost further, hence the importance of having a fast mixing rate transition matrix. When a Worker suspects that some other Worker on a Hive is faulty for too long, he first contacts a dedicated server, called Keeper, who decides the new Hive membership. The suspecting Worker then individually decides based on previous aggregate information if he should inject new fragments regarding the file into the network if he believes durability is at risk. Whenever a Worker who has been temporarily offline reconnects, he first contacts the master to know what \textit{fileId} is he is no longer responsible for maintaining, allowing him to free space. They also perform this procedure whenever they do not receive a fragment or a ping regarding a \textit{fileId} for a sufficiently long period.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN EVALUATION METHODOLOGY %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation Methodology}\label{sec:methodology}
We will first evaluate the performance of different transition matrix generating algorithms by comparing their theoretical fast mixing property against the number of discrete time steps the matrices take to converge on a simple, cycle-based, locally ran, custom-built simulator. We will select one of these algorithms based on how fast convergence occurs, computing time vs. space complexity, and the possibility of allowing each peer to update their transition matrix autonomously when faced with transient neighbor faults. After this initial step, we will implement a reactive replica-control policy using erasure-coding at each peer and test the durability of files on our custom-built simulator, where we have perfect failure detectors. We are concerned with discovering a good trade-off between storage overhead and durability by using different (n, k) values without caring for any networking aspects. Finally, we will measure bandwidth consumption as well as the robustness of our solution under more realistic simulations by either implement our protocol on well-known P2P Simulation test-bed such has PeerfactSim.KOM\cite{peerfact}, or if time permits it, deploy proof of concept applications on the university's computer cluster for distributed testing.

@@

%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN WORK SCHEDULE %
%%%%%%%%%%%%%%%%%%%%%%%
\section{Work Schedule}\label{sec:workschedule}
%%%%%%%%%%%%%%%%%%%%%
% END WORK SCHEDULE %
%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%
% BEGIN REFERENCES %
%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{splncs04}
\bibliography{bibliography}
%%%%%%%%%%%%%%%%%%
% END REFERENCES %
%%%%%%%%%%%%%%%%%%

>>>>

\begin{itemize}
    \item Identify a steady-state, $Q$, providing optimal robustness and load-balancing
    \item Identify a transition matrix, $M$, that allows peers to achieve $Q$ collectively
    \item Optimize $M$ with respect to peer connectivity degree
    \item Optimize updates to $M$ as overlay topology changes over time
    \item Implement a replication model that exploits the properties of the $M$
    \item Evaluate solution performance in a simulation or on in real life scenario
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
% END INTRODUCTION %
%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{images/gfs_paper.png}
\caption{GFS Architecture}
\label{fig:gfs_arch}
\end{figure}

@@

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.45\textwidth]{images/cephfs.PNG}
    \caption{Ceph File System Architecture \cite{configure-ceph}}
    \label{fig:ceph_arch}
\end{figure}

\newpage\section{P2P Overlays}\label{sec:p2p-overlays}

\begin{figure}[ht]
    \centering
    \subfigure{\label{fig:iplayer}\includegraphics[width=0.35\textwidth, clip=true]{images/ip-layer.png}}
    \hspace{0.2\textwidth}
    \subfigure{\label{fig:overlay}\includegraphics[width=0.35\textwidth, clip=true]{images/overlay-layer.png}}
    \caption{P2P Network Diagrams: IP layer \textit{vs.} One possible overlay abstraction}
    \label{fig:network_diagrams}
\end{figure}

@@

\newpage\section{Membership Management Frameworks}\label{sec:membership-management}
Gossip-based protocols are well known for their usability to solve problems like database replication, failure detection, and resource-monitoring. We present two frameworks used for membership management that ensure peers keep up to date partial views of large-scale, dynamic networks, built on unstructured overlays using such protocols. Just like we mentioned in section \ref{sec:p2p-overlays}, keeping a view of all peers in the system is often not a solution. Another fundamental issue in P2P is how to avoid partitioning. The following alternatives are both scalable, robust, and decentralized solutions to both these problems. They gift peers with autonomous capabilities regarding self-organization and overlay healing.

@@

\newpage\section{Cloud-Assisted P2P Networks}\label{sec:ca-networks}

@@

\newpage\section{Data Redundancy}\label{sec:dataredundancy}
In DFS, not losing data is of the utmost importance. Particularly, peer availability in P2P networks is ever-changing and not guaranteed, and thus, availability, and even durability of files are at risk. A \textbf{\textit{durable file}} is one that, once put into the system, is not lost due to failures, whereas availability means that a user can access it promptly. Durability is accomplished by employing file or block-level \textbf{\textit{replication}} or alternatively, using \textbf{\textit{erasure coding}} techniques (\textit{coding} for short). Hybrid approaches are also possible. Replication has the advantage of having a straightforward and less resource-intensive recovery model, on the other hand, coding has a better \textit{expansion factor}, i.e., can provide the same or even increased durability at a lower storage overhead. Because our proposal focuses on durability of files in P2P networks built on top of personal computers, where resources may be limited, we survey only coding, mentioning raised concerns and inherent disadvantages of this technique.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN SOLUTION PROPOSAL %
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Solution Proposal}\label{sec:proposal}
In this section, we introduce a prototype architecture that targets the use case that inspired to explore the theme of this thesis. We explain it by decomposition.

\subsection{Use Case}
Fundamentally, the idea is that a company, Hives, with dedicated web servers called Hiveminds, provides a platform for users to sell storage belonging to their personal computers directly to other users. Clients pay a monthly fee directly to Hives based on how much storage space they need. Hiveminds then puts together a set of sellers, which constitute a Hive, that serves adequate storage space to the clients. Sellers are selected according to the reputation they earn over time, based on their reliability, with a bias towards more reputable sellers. The collected fee is distributed proportionately among Hive members according to their overall space contribution and up time, calculated weekly, for example. The company's revenue comes from keeping a small percentage of the monthly fees. If the client requires continuous availability, rather than durability, he may pay a premium, in which case, inspired on the related work mentioned in section \ref{sec:ca-networks}, cloud storage may be used to help the Hive in meeting the required service. Seller shares are reduced if cloud storage is used at any point.

@@

\subsection{Hiveminds - Master Servers}
The Hivemind servers are the backbone of the system; they have to handle registration, login, subscription and payment procedures. When a client wishes to subscribe to a monthly plan, he contacts one Hivemind by accessing his desktop application, and the Hivemind then sets a \textit{usedSpace} variable indicating how much data the Client uploaded into the system to zero. This variable is updated according to the Client's future requests. In Hives File System (Hives), users' files are uniquely identified and have no hierarchy. Hiveminds keep a mapping between the \textit{fileId} and the Hive\footnote{Set of peers S responsible for holding one or more files from one or more clients} that is responsible for storing them. Hive membership is dynamic, and a mapping between the \textit{HiveId} and the Hive members (Workers) is also kept. Note that it is not relevant to know which fragments each Worker in the Hive has; this is important because it diminishes the number of messages sent to Hiveminds and also reduces space overhead required to keep track of file locations. When a Hivemind receives a file upload request from a Client, it finds or creates a suitable Hive to host the data and calculates, if needed, the transition matrix used for PSG within it, before returning all relevant data to the Client, including Workers' addresses. When the Hivemind receives a delete request, it drops all mapping data regarding the respective \textit{fileId}.

\subsection{Clients}
After registration, logging in, and subscription procedures, Clients, can perform four main procedures, which we name and implement using HTTP semantics. In all cases, the Client first retrieves the Hive Workers' addresses. In a GET request, the Client downloads enough fragments from the Hive Workers and reconstructs the file himself. In a DELETE request, the Client first asks the Hivemind to stop tracking efforts, and once he receives an acknowledge, he broadcasts in a best-effort manner a DELETE message to any online worker within that file's Hive. In a POST operation, the Client first obtains permission to upload from the Hivemind; after that, he divides his file into parts, applies an erasure coding algorithm over each them, and uploads the encoded fragments to the workers at random. PUT updates existing files and operate similarly to POST, but whenever possible, tries to send the new version to the Hive that contains the previous \textit{versionId}.

\subsection{Workers - Sellers}
Apart from adequately dealing with clients' requests, Workers organize themselves in a P2P fashion within each Hive. One Worker can be in zero or more Hives, and within each Hive, they continuously redistribute encoded file part fragments according to the transition matrix that is calculated by the Hivemind and delivered to them by a Client. In the long run, we expect that a Hive achieves the desired fragment distribution that fairly distributes load while ensuring optimal durability. Constantly exchanging fragments is expected to cause a bandwidth overhead but has the advantage that Clients can upload their fragments to any online peer or peers in the Hive instead of one single entity that may go offline during the upload. Workers will then redistribute the fragment using simple routing rules. This approach does not sacrifice fast lookup operations due to the structured nature of the system using a Hivemind that behaves somewhat like a BitTorrent\cite{bittorrent} tracker. The never-ending fragments exchange between peers also facilitates the timely detection of non-transient faults in the Hive, which helps to ensure file durability. Workers piggyback aggregate information to the fragments, including but not limited to, how many parts of a file they have, allowing us to implement a hybrid between a reactive and a proactive system concerning replica control, in hopes of minimizing bandwidth burst on top of already high bandwidth requirements. Fragment exchange may eventually be replaced with ping-like messages to minimize bandwidth cost further, hence the importance of having a fast mixing transition matrix. When a Worker suspects that some other Worker on a Hive is faulty for too long, he first contacts a dedicated server, called Keeper, who decides the new Hive membership. The suspecting Worker then individually decides based on previous aggregate information if he should inject new fragments regarding the file into the network if he believes durability is at risk. Whenever a Worker who has been temporarily offline reconnects, he first contacts the master to know what \textit{fileId} is he is no longer responsible for maintaining, allowing him to free space. They also perform this procedure whenever they do not receive a fragment or a ping regarding a \textit{fileId} for a sufficiently long period.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN EVALUATION METHODOLOGY %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Evaluation Methodology}\label{sec:methodology}
We will first evaluate the performance of different transition matrix generating algorithms by comparing their theoretical fast mixing property against the number of discrete time steps the matrices take to converge on a simple, cycle-based, locally ran, custom-built simulator. We will select one of these algorithms based on how fast convergence occurs, computing time vs. space complexity, and the possibility of allowing each peer to update their transition matrix autonomously when faced with transient neighbor faults. After this initial step, we will implement a reactive replica-control policy using erasure-coding at each peer and test the durability of files on our custom-built simulator, where we have perfect failure detectors. We are concerned with discovering a good trade-off between storage overhead and durability by using different (n, k) values without caring for any networking aspects. Finally, we will measure bandwidth consumption as well as the robustness of our solution under more realistic simulations by either implement our protocol on well-known P2P Simulation test-bed such has PeerfactSim.KOM\cite{peerfact}, or if time permits it, deploy proof of concept applications on the university's computer cluster for distributed testing.

@@

%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN WORK SCHEDULE %
%%%%%%%%%%%%%%%%%%%%%%%
\section{Work Schedule}\label{sec:workschedule}
\begin{itemize}
    \item January 6 to January 12: Implement PSG on basic custom made simulator;
    \item January 13 to January 19: Perform preliminary tests;
    \item January 20 to January 26: Improve fast mixing property;
    \item January 27 to February 2: Write a paper describing initial results;
    \item February 3 to February 9: Implement erasure coding algorithm
    \item February 10 to February 16: Holiday Period
    \item February 17 to February 23: Evaluate and improve erasure coding algorithm
    \item February 24 to March 29: Attempt to port solution to PeerfactSim.KOM
    \item March 30 1 to April 12: Preliminary tests and possible bug fixes
    \item April 13 to April 26: Perform the extensive experimental evaluation
    \item April 27 to June 15: Write draft for MSc dissertation
    \item June 16 to August 30: Attempt to deploy proof of concept application
    \item August 31 to September 27: Finish writing and deliver MSc dissertation
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%
% END WORK SCHEDULE %
%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%
% BEGIN REFERENCES %
%%%%%%%%%%%%%%%%%%%%
\newpage
\bibliographystyle{splncs04}
\bibliography{bibliography}
%%%%%%%%%%%%%%%%%%
% END REFERENCES %
%%%%%%%%%%%%%%%%%%
